---
title: "Human activity recognation"
output: html_document
---

Below we use random forrest machine learning algorithm to classify human acitivities based on accelerometer measurements on various parts of bodies for 6 participants.

## Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset). 

## Executive summary

Based on the data provided from accelerometer measurements, we performed machine training to classify the types of motion. The training data is split into a training and validation set (75% and 25% of the training data, respectively). For all the 160 variables given in the data set, we performed preprocessing by keeping only the accelerometer measurements, and removing those with 80% data missing. A principle component analysis is then used to remove excessive measurements that are highly correlated the others. 25 principle components remains after preproocessing, and With the data trained with a random forest model, 100% accuracy is achieved on the training set. Using this same model a 2.5% out of sample error is found in the validation set. Finally, this model predicts the outcome from the testing set to be "B A B A A E D B A A B C B A E E A B B B", on the 20 sets of measurements.



## Getting and cleaning data
Training data: 
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data: 
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

### Downloading data:

```{r, echo = TRUE}
#download.file(url = "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", destfile = "har-training.csv")
#download.file(url = "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", destfile = "har-testing.csv")
har_train <- read.csv('har-training.csv')
har_test <- read.csv('har-testing.csv')
```

### Initial examination of data:
```{r, echo = TRUE}
dim(har_train)
names(har_train)
```
The goal is to build a model based on the 5 sensor readings attached to the patient to predict the variable classe. We therefore perform the following to pre-clean the data:
* Use only the 5 sensor readings as predictors, which are labeled "belt", "arm", "forearm", and "dumbbell", respectively.
* Of all the readings, we's also like to remove variables which have more than 80% empty/NA entries.

### Cleaning data (removing variables with all zero entries):

```{r, echo = TRUE}
sensorInd <- grepl("classe|belt|[^(fore)]arm|dumbbell|forearm", names(har_train))
rmColInd <- colSums(is.na(har_train) | har_train == "")/dim(har_train)[1] > 0.8
har_train <- har_train[,sensorInd & !rmColInd]
har_test <-har_test[,sensorInd & !rmColInd]
dim(har_train)
names(har_train)
```

## Preprocessing

Loading all libraies needed
```{r, echo =TRUE}
library(caret)
#library(ggplot2)
#library(lattice)
#library(kernlab)
#library(randomForest)
```

Partitioning training data set into training (data_train) and validation (data_val) set.

```{r, echo = TRUE}
set.seed(6639)
index_train <- createDataPartition(har_train$classe,p=.75,list=FALSE)
data_train <- har_train[index_train,]
data_val <- har_train[-index_train,]
```

Preprocessing using the PCA method:
```{r, echo = TRUE}
preProc <- preProcess(data_train[,names(data_train)!="classe"], method = "pca")
predictor_trainPC <- predict(preProc, data_train[,names(data_train)!="classe"])
data_trainPC <- data.frame(classe = data_train$classe,predictor_trainPC)
# examining the number of principle components used
dim(data_trainPC)
```

## Training

Now we're ready to perform training with random forest methods
```{r, echo = TRUE}
# training with preprocessed principle components
modelTrain <- train(classe ~ ., method = "rf", data = data_trainPC)
confusionMatrix(data_train$classe, predict(modelTrain, data_trainPC))
```
As revealed from the output, the 25 predictors as selected are sufficient to determine the training set with 100% accuracy.

## Validation

Now try the model on the validation set:
```{r, echo = TRUE}
predictor_valPC <- predict(preProc, data_val[,names(data_val)!="classe"])
data_valPC <- data.frame(classe = data_val$classe,predictor_valPC)
confusionMatrix(data_val$classe, predict(modelTrain, data_valPC))
```
The out-of-sample error is 0.025.

## Testing

Using the model, we're now ready to predict the results in the testing data set:
```{r, echo = TRUE}
predictor_testPC <- predict(preProc, har_test[,names(har_test)!="problem_id"])
predResult <- predict(modelTrain, predictor_testPC)
print(predResult)
```
## Making files for submission:
```{r, echo = TRUE}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
pml_write_files(predResult)
```


